Windowed least-squares model reduction for dynamical systems

### 2019-10-17 numerical_experiments.tex
minor changes

\section{Numerical experiments}\label{sec:numerical_experiments}
This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity. In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs. The purpose of the numerical experiments is to assess the impact on minimizing the residual over an arbitrarily sized time window on the solution accuracy. We additionally assess the impact of the time step, and time scheme on \methodAcronym. In both experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$. It is also noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs. Future state prediction and prediction at new parameter instances are not considered as these problems introduce factors that confound the solution accuracy with the solution methodology (e.g., accuracy of the basis).

@@

\item \textit{LSPG ROM:} The LSPG ROM is built on top of the FOM discretization leveraging the CN time scheme as previously described. Unless noted otherwise, the LSPG ROM uses a constant time step of $\Delta t = 0.002$. The nonlinear least-squares problem arising at each time instance is solved via the Gauss--Newton method. All Jacobians are computed via automatic differentiation~\cite{adolc}. The Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$. 
\item \textit{\methodAcronymROM:} \methodAcronymROMs\ are solved via the direct and indirect methods with two different time-stepping schemes: 
\begin{itemize}
\item Direct method: We consider \methodAcronymROMs\ solved via the direct method for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time step of $\Delta t = 0.0005$. The nonlinear least-squares problem 
arising over each window is solved via the Gauss--Newton method. All Jacobians are again computed via automatic differentiation, and the Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $1e-4$. Critically, it is noted that the Jacobian 
matrix over a window is constructed by assembling local (dense) Jacobians. The global Jacobian over a single window is then stored in a compressed sparse row format. Uniform quadrature weights are used for evaluating the integral in~\eqref{eq:obj_gen_slab}.

\item Indirect method: We consider two \methodAcronymROMs\ solved via the indirect method. The first uses the same CN discretization at at time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$. The coupled two-point boundary 
problem is solved via the forward--backward sweep method. The action of the Jacobian transpose on vectors is computed via automatic differentiation. 
\end{itemize}
\end{itemize}

>>>>

\section{Numerical experiments}\label{sec:numerical_experiments}
This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity. In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs. The purpose of the numerical experiments is to assess the impact of minimizing the residual over an arbitrarily sized time window on the solution accuracy. We additionally assess the impact of the time step and time scheme on \methodAcronym. In both experiments, the spatial basis is equivalent for each time window, e.g., $\basisspaceArg{n} \equiv \basisspace$, $n=1,\ldots,\nslabs$. It is also noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs. Future state prediction and prediction at new parameter instances are not considered as these problems introduce factors that confound the solution accuracy with the solution methodology (e.g., accuracy of the basis).

@@

\item \textit{LSPG ROM:} The LSPG ROM is built on top of the FOM discretization leveraging the CN time scheme as previously described. Unless noted otherwise, the LSPG ROM uses a constant time step of $\Delta t = 0.002$. The nonlinear least-squares problem arising at each time instance is solved via the Gauss--Newton method. The linear least-squares problems arising at each Gauss--Newton iteration are solved via the normal equations. All Jacobians are computed via automatic differentiation~\cite{adolc}. The Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $10^{-4}$. 
\item \textit{\methodAcronymROM:} \methodAcronymROMs\ are solved via the direct and indirect methods with two different time-stepping schemes: 
\begin{itemize}
\item Direct method: We consider \methodAcronymROMs\ solved via the direct method for both the same CN discretization employed in the FOM and LSPG, as well as for the second-order explicit Adams Bashforth (AB2) discretization using a constant time step of $\Delta t = 0.0005$. The nonlinear least-squares problem 
arising over each window is solved via the Gauss--Newton method. The linear least-squares problems arising at each Gauss--Newton iteration is solved via the normal equations. All Jacobians are again computed via automatic differentiation, and the Gauss--Newton algorithm is deemed converged when the gradient norm is lower than $10^{-4}$ (i.e., we use the parameter $\epsilon = 10^{-4}$ in Algorithm~\ref{alg:colloc_gn}). Critically, it is noted that the Jacobian 
matrix over a window is constructed by assembling local (dense) Jacobians. The global Jacobian over a single window is then stored in a compressed sparse row format. Uniform quadrature weights are used for evaluating the integral in~\eqref{eq:obj_gen_slab}.

\item Indirect method: We consider two \methodAcronymROMs\ solved via the indirect method. The first uses the same CN discretization at at time step of $\Delta t = 0.002$, while the second uses the AB2 discretization using a time step of $\Delta t = 0.0005$. The coupled two-point boundary 
problem is solved via the forward--backward sweep method. The action of the Jacobian transpose on vectors is computed via automatic differentiation. We use parameters $\epsilon = 10^{-6}$, $\fbsmGrowth = 1.1$, and $\fbsmDecay=2$ in Algorithm~\ref{alg:st_iter}. 
\end{itemize}
\end{itemize}
